The promise of community-driven preprints in ecology and evolution

### 2024-06-03 ms/ms.qmd
Up to generational with edits

A significant barrier to the communication of research findings is due to the fact that they are primarily communicated in English [@amano2016languages; @amano2013four; @amano2021ten]. Communication of research through a single language has major consequences for the global distribution of knowledge resulting in knowledge gaps across some of the most biodiverse and threatened regions in the world [@amano2023role; @zenni2023multi]. Such gaps also have flow on effects for research syntheses and meta-analyses because they create a distorted picture of our knowledge base that can lead to bias and inadequate sampling that affects policy development and decision-making [@hannah2024language; @zenni2023multi; @white2021geographical; @konno2020ignoring].

*EcoEvoRxiv* is the only preprint server to date breaking down language barriers to scientific communication; currently accepting English, Spanish, Portuguese and French manuscripts. Such initiatives are incredibly important if we are to begin filling global voids of scientific knowledge. However, multilingual initiatives have been slow to take off on *EcoEvoRxiv*, with only a few Spanish preprints and a single Portuguese preprint posted to date. Part of the challenge in getting authors to submit non-English preprints is the lack of awareness of *EcoEvoRxiv* in non-English speaking countries and the fact that it accepts preprints in different languages, cultural differences in the perception of preprints, and a strong reliance on traditional publishing models which typically mandate publishing in English.

>>>>

A significant barrier to the communication of research findings is the fact that they are primarily communicated in English [@amano2016languages; @amano2013four; @amano2021ten]. Communication of research through a single language has major consequences for the global distribution of knowledge, resulting in knowledge gaps across some of the most biodiverse and threatened regions in the world [@amano2023role; @zenni2023multi]. Such gaps also impact research syntheses and meta-analyses because they create a distorted picture of our knowledge base leading to bias and inadequate sampling that can affect policy development and decision-making [@hannah2024language; @zenni2023multi; @white2021geographical; @konno2020ignoring].

*EcoEvoRxiv* is the only preprint server to date that breaks down language barriers to scientific communication by accepting English, Spanish, Portuguese and French articles. *EcoEvoRxiv* plans to expand to other languages as new non-English editors become available. Such initiatives are incredibly important if we are to begin filling global voids of scientific knowledge. However, multilingual initiatives have been slow to take off on *EcoEvoRxiv*, with only a few Spanish articles, and a single Portuguese article, posted since starting to accept non-English articles in 2023. Part of the challenge in getting authors to submit non-English articles is the lack of awareness of *EcoEvoRxiv* in non-English speaking countries, cultural differences in the perception of articles, and a strong reliance on traditional publishing models which typically mandate publishing in English [@arenas2024academic].
